The ChEMBL database as Linked Open Data

### 2013-03-26 article.tex
reword "not changed much" sentence
reword are to is, as the subject of the sentence seems to be the ontology and not the concepts

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF triples using a set of PHP scripts. The scripts use SQL to
query the database and output triples using print statements. Custom PHP scripts have been used
as a light-weight solution to provide a Linked Data API via a Apache webserver, as well as to
create data files with triples. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
Group~\cite{Marshall2012}. This process has not changed much since its original description.

However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
in the relational database, and are not formally defined. The current version, instead, uses community-proposed
ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
The new triple dataset uses standard ontologies
including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.

>>>>

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL data dump provided by ChEMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF triples using a set of PHP scripts. The scripts use SQL to
query the database and output triples using print statements. Custom PHP scripts have been used
as a light-weight solution to provide a Linked Data API via a Apache webserver, as well as to
create data files with triples. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences Interest
Group~\cite{Marshall2012}. This process is similar to its original description.

However, the previous ChEMBL-RDF conversion used custom predicates and classes to represent RDF triples, defining
an ad-hoc ontology~\cite{Willighagen2011}. This implicit ontology reflects the concepts expressed
in the relational database, and is not formally defined. The current version, instead, uses community-proposed
ontologies, making the RDF more interoperable with the RDF offerings of different databases. 
The new triple dataset uses standard ontologies
including the Bibliography Ontology~\cite{Giasson2011} and the Citation Typing Ontology (CiTO)~\cite{Shotton2010} for literature
references and citations, and domain ontologies such the Protein Ontology\cite{Natale2011} and the Chemical Information
Ontology\cite{Hastings2011}. Throughout this paper various prefixes are used to denote different ontology 
namespaces to simplify the RDF examples. These prefixes are outlined in Table~\ref{namespaces}.
